Measurement of total hadronic differential cross sections in the LArIAT experiment

### 2018-04-28 Chapter-2/2-LArTPCs.tex
s

\subsection{Signal Processing and Event Reconstruction}
In this section we illustrate the processing and reconstruction chain of the TPC signals, from the pulses on the sense wire to the construction of three dimensional objects with associated calorimetry.\\

\textbf{Deconvolution.} Induction and collection planes have different field responses, given the different nature of the signals on these planes: the wires on the induction planes see inductive signal of the drifting charge, while the wires on the collection planes see the current derived from the charge entering the conductor. Thus, signals on the induction plane are bi-polar pulse and signal on the collection plane are unipolar pulses, see Figure \textcolor{red}{ADD PULSES FIGURE}. The first step in signal processing is deconvolution, that is a series of off-line algorithms geared towards undoing the detector effects. The result of the deconvolution step is  the production of  a comparable set waveforms on all planes presenting unipolar, approximately gaussian-like pulses. Signal from all planes are treated on equal footage beyond this point. Some LArTPC apply noise filtering in the frequency domain just after the deconvolution to clean up wire cross talk. Since signals from the LArIAT TPC are extremely clean, noise filtering is not necessary.\\

\textbf{Hit Reconstruction.} The second stage of the signal processing is the reconstruction of hits, indicating an energy deposition in the detector.  A peak finder scans the deconvolved TPC waveforms for each wire on the whole readout time looking for spikes  above  the waveform's baseline. It then fits these peaks with gaussian shapes and stores the fit parameters such as the quality of the fit, the peak time, height and area under the gaussian fit. The information resulting fron this process a single spike form a single reconstructed ``hit".
The next steps in the event reconstruction chain will then decide if rejecting hits with poor fits.
It is important to notice how the height and width of the hit depend on the topology of the event: for example, a particle running  parallel to the wire planes will leave a series of sharp hits on many consecutive wires, while a particle traveling towards the planes will leave a long, wide hit on very few wires. The height of the hits and their integral is proportional to the charge collected on the wire, so it depends on the particle type.

The event reconstruction chain uses collection of hits to form more complex objects associated with the particles in the detector. The development of different approaches to accomplish this task is an extremely hot topic in LArTPC event reconstruction which spans from more traditional approaches such as line-clustering  \cite{MicroBooNE} \textcolor{red}{argoneut paper} to the use of machine learning tools \cite{1748-0221-12-03-P03011}. Generally speaking, the scope of hit clustering and event reconstruction to provide shower-like or track like-objects with an associated energy reconstruction. This is because different particles have different topology in the detector -- electrons and photon create electromagnetic showers,  resulting in shower-like topologies, while muons and hadrons  leave track-like signals.  For the scope of these thesis, we will describe only LArIAT's approach to track reconstruction even if we recognize the breath of LArTPC event reconstruction is much wider. We are interested in the reconstruction of pions and kaons in the active volume, whose topology is track-like.\\

\textbf{2D Clustering Reconstruction.} 
The LArIAT reconstruction of track-like objects starts by clustering hits on the collection and induction planes separately with the use of the TrajCluster clustering package\cite{Bruce} \textcolor{red}{Bruce}. 
TrajCluster looks for a collection of hits in the wire-time 2D space which can be described with a line-like 2D trajectory. TrajCluster reconstructs trajectories by adding trajectory points to the leading edge of the trajectory while stepping through the 2D space of hits. Several factors determine whether a hit is added to the trajectory, including but not limited to
\begin{enumerate}
\item the goodness of the fit of the single hit,
\item the charge of the hit compared to the average charge and RMS of the hits already forming the trajectory,
\item the goodness of trajectory fit with and without the hit addition,
\item the angle between the two lines formed by the collection of hits before and after the considered hit in the trajectory.
\end{enumerate}

\textbf{3D TrackReconstruction.} The 3D tracking set of algorithms uses clusters close in time on the induction and collection planes as starting point to form a 3D track. Firstly, it construct a tentative 3D trajectory using the edges of the clusters. Then, it  projected back the tentative trajectory on to the planes and adjusts the parameters of the 3D track fit such that they minimize the distance between the fit projections and the track hits in all wire planes simultaneously.  The track algorithm can use multiple clusters in one plane, but it can never break them in smaller groups of hits. This algorithm was first developed for the ICARUS collaboration\cite{Antonello2013}.

@@

\begin{figure}[hbpt]
\centering
\includegraphics[width=6.5in]{Chapter-2/Images/PDKExperimentalLImit.png}
\caption{Proton decay lifetime limits from passed and future experiments.}
\label{fig:PDKExperimentalLImit}
\end{figure}

>>>>

\subsection{Signal Processing and Event Reconstruction}
In this section we illustrate the processing and reconstruction chain of the TPC signals, from the pulses on the sense wire to the construction of three dimensional objects with associated calorimetry. Different experiments can chose different software packages for their off line signal processing and event reconstruction, but  a popular choice for  US based  LArTPCs is LArSoft \cite{EricFChurck}. Based on the Art framework \cite{Green:2012gv}, LArSoft is an event-based toolkit to perform simulation, analysis and reconstruction of LArTPCs events.\\

LArTPC signal processing develops in several consecutive stages that we summarize here in the following categories: \emph{Deconvolution}, \emph{Hit Reconstruction}, \emph{2D Clustering}, \emph{3D Tracking}, \emph{Calorimetry Reconstruction}.  A visualization of the signal processing workflow is shown in figure \ref{fig:SignalProc}.\\

\begin{figure}[hbpt]
\centering
\includegraphics[width=\textwidth]{Chapter-2/Images/SignalProc.jpg}
\caption{A scheme of a typical signal processing workflow in LArSoft.}
\label{fig:SignalProc}
\end{figure}

\textbf{Deconvolution.} Induction and collection planes have different field responses, given the different nature of the signals on these planes: the wires on the induction planes see inductive signal of the drifting charge, while the wires on the collection planes see the current derived from the charge entering the conductor. Thus, signals on the induction plane are bi-polar pulse and signal on the collection plane are unipolar pulses, see Figure \ref{fig:SignalProc} panel a. The first step in signal processing is deconvolution, that is a series of off-line algorithms geared towards undoing the detector effects. The result of the deconvolution step is  the production of  a comparable set waveforms on all planes presenting unipolar, approximately gaussian-like pulses (Figure \ref{fig:SignalProc} panel b). Signal from all planes are treated on equal footage beyond this point. Some LArTPC apply noise filtering in the frequency domain just after the deconvolution to clean up wire cross talk. Since signals from the LArIAT TPC are extremely clean, noise filtering is not necessary.\\

\textbf{Hit Reconstruction.} The second stage of the signal processing is the reconstruction of hits, indicating an energy deposition in the detector.  A peak finder scans the deconvolved TPC waveforms for each wire on the whole readout time looking for spikes  above  the waveform's baseline. It then fits these peaks with gaussian shapes and stores the fit parameters such as the quality of the fit, the peak time, height and area under the gaussian fit. The information resulting fron this process a single spike form a single reconstructed ``hit".
The next steps in the event reconstruction chain will then decide if rejecting hits with poor fits.
It is important to notice how the height and width of the hit depend on the topology of the event: for example, a particle running  parallel to the wire planes will leave a series of sharp hits on many consecutive wires, while a particle traveling towards the planes will leave a long, wide hit on very few wires. The height of the hits and their integral is proportional to the charge collected on the wire, so it depends on the particle type.\\

The event reconstruction chain uses collection of hits to form more complex objects associated with the particles in the detector. The development of different approaches to accomplish this task is an extremely hot topic in LArTPC event reconstruction which spans from more traditional approaches such as line-clustering  \cite{Barker2011} to the use of machine learning tools \cite{1748-0221-12-03-P03011}. Generally speaking, the scope of hit clustering and event reconstruction to provide shower-like or track like-objects with an associated energy reconstruction. This is because different particles have different topology in the detector -- electrons and photon create electromagnetic showers,  resulting in shower-like topologies, while muons and hadrons  leave track-like signals.  For the scope of these thesis, we will describe only LArIAT's approach to track reconstruction even if we recognize the breath of LArTPC event reconstruction is much wider. We are interested in the reconstruction of pions and kaons in the active volume, whose topology is track-like.\\

\textbf{2D Clustering Reconstruction.} 
The LArIAT reconstruction of track-like objects starts by clustering hits on the collection and induction planes separately with the use of the TrajCluster clustering package\cite{Baller2016}. 
TrajCluster looks for a collection of hits in the wire-time 2D space which can be described with a line-like 2D trajectory. TrajCluster reconstructs trajectories by adding trajectory points to the leading edge of the trajectory while stepping through the 2D space of hits. Several factors determine whether a hit is added to the trajectory, including but not limited to
\begin{enumerate}
\item the goodness of the fit of the single hit,
\item the charge of the hit compared to the average charge and RMS of the hits already forming the trajectory,
\item the goodness of trajectory fit with and without the hit addition,
\item the angle between the two lines formed by the collection of hits before and after the considered hit in the trajectory.
\end{enumerate}

\textbf{3D Tracking.} The 3D tracking set of algorithms uses clusters close in time on the induction and collection planes as starting point to form a 3D track. Firstly, it construct a tentative 3D trajectory using the edges of the clusters. Then, it  projected back the tentative trajectory on to the planes and adjusts the parameters of the 3D track fit such that they minimize the distance between the fit projections and the track hits in all wire planes simultaneously.  The track algorithm can use multiple clusters in one plane, but it can never break them in smaller groups of hits. This algorithm was first developed for the ICARUS collaboration\cite{Antonello2013}.\\

\textbf{Calorimetry.} The last step in the event reconstruction chain is to assign calorimetric information to the track (or shower) objects. Calorimetry is performed separately on the different planes. A multi-step procedure is needed to retrieve the energy deposited in the TPC  from the charge seen by the wires.
For each hit associated with the track object, the calorimetry algorithms calculate the charge seen on every wire using the area underneath the gaussian fit; then they correct this raw charge by the electron life time, the electronic noise on the considered wire and the recombination effect. Lastly an overall calibration of the energy, explained in detail in section \ref{ch:energyCalibration}, is applied and the calorimetric information for the given track is assigned.
Even if calorimetry is done in 2D, it benefits from the 3D tracking information; typical information available after the calorimetric reconstruction are the total energy deposited by the particle and its stopping power $dE/dx$ at each ``track pitch", i.e. at each 2D projection on the wire plane of the 3D trajectory.

@@

\begin{figure}[hbpt]
\centering
\includegraphics[width=\textwidth]{Chapter-2/Images/PDKExperimentalLImit.png}
\caption{Proton decay lifetime limits from passed and future experiments.}
\label{fig:PDKExperimentalLImit}
\end{figure}
